A Graph Theoretic Approach to Problem Formulation for Multidisciplinary Design Analysis and Optimization

### 2013-09-04 reviewer_response.tex

\rev{If PSG is covered even to a small 
extent (like explained in first comment) I would suggest or like to see the 
PSG equivalent of problem formulation shown in Figure 2 with XDSM.}

See new Fig. 8 and Fig. 13

\rev{Is it really adding a lot of value to have graphs like MCG or FPG which 
are very explicit in nature and can become difficult to read}

Their value is not derived from their ability to represent the data in a human readable 
form, but in it's power as a machine readable one. Section 5.3 introduces an 
algorithm for testing if graphs are in fact reduced down to fundamental problems and 
to reduce them to one if they are not. As problems grow larger and more complex, 
this kind of algorithm becomes ever more valuable.

It could be implemented into a framework or stand alone tool to help teams develop the
specifics of a design problem before they started working to code the modules and such. 
Section 6 shows that for even simple problems, the graph (as viewed by an algorithm) is 
non-trivial. It also shows that there are multiple ways to solve problems. As 
problems grow even a small amount from the size given in the example it's clear that 
algorithmic analysis useful. The graph syntax derives its usefulness from 
the effectiveness of describing the graph to a computer.

\rev{Can the graph be compressed in a logical way to reduce the number of nodes 
and their interconnections while still conveying important and critical 
information for that stage (MCG stage or FPG stage) ? Reducing cycles is 
discussed but the graph may still look complex.

Looking at these graphs (for example Fig. 8), a node for example Z2 
appears several times. It may be useful to figure out holes and collision, 
but once it is resolved, can those nodes be consolidated to provide simpler 
visualization ?}

There are a number of ways the graph could be simplified down to make it more 
human readable, but this was not the primary goal of the work. You could combine 
variables or remove them all together and just represent component to component 
connections like a pure DSM.

It should be possible to convert the entire PSG graph into the that used by the 
data-connectivity part of an XDSM. You would have all the information in a more 
compact and visually insightful form. You could always convert that back into the 
more verbose syntax here as well. We would argue that the XDSM is more human readable, 
but this syntax is a closer representation of how a framework would view the problem.

Given our goals, the verbose nature of the graphs is a necessary side effect.

@@

\rev{As the problem size increases, how will it affect the size and complexity 
of such graphs.}

The answer to this question is not trivial, as it depends on a number of factors 
such as the number of analyses and the number of variables in those analyses. 
However, speaking roughly, you could measure the complexity of a graph by the total 
number of nodes and edges.

So the number of nodes will grow roughly linearly with the number of analysis tools 
and the number of variables. The number of fixed edges will grow linearly with the number of variables as well. The number of connection edges depends entirely on the specifics of the problem, but since any variable can only have a finite number of incoming edges (usually 1, but maybe 2 or 3 in multi-fidelity case) you can avoid an unbounded 
growth there. So very roughly speaking, the graph will grow O(N+M) where N is the number 
of variables and M is the number of analyses.

If you assume large problems have hundreds of components and millions of variables, 
you're talking about graphs with a few million nodes and edges. These are very large, but 
not outside the relm of what is manageable by todays graph libraries.

@@

\rev{
I will assume that creation of such a graph and interconnections will be 
expected to be done by some kind of software tool or wizard. A brief 
discussion about why this MCG, FPG and PSG idea is better to lead to such a 
system can be useful. Although this itself can be a separate research and 
paper, but why this idea can provide a better foundation can be valuable.
}

See the new section 2.4 for a short section describing our envisioned usage 
of the graph language.

[TODO: Brian and David can flush this out better than me...]

@@

\rev{Minor comments:
Section 1 Line 46: I think other commercial tools such as iSight, 
modeFRONTIER should also be mentioned since their usage is more widespread 
compared to OpenMDAO.}

Agreed. Done and incorporated into Section 1.

\rev{Page 13 line 33 column 2: Typo : “Step (D) now PROCEDEDS … “}

fixed.

\rev{Page 16 line 38 
column 2: Typo : “… formulations involving MULTIPLES analysis ….”}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\rev{Line 10, 1st column, Page 10:  This equation (missing a number) has a LOT of 
unions, variables, etc...in it.  It would help tremendously to provide a 
verbal description of at least one of these equations to help with 
translation into the real world.  This would be similar to what you did 
following the steps (1) to (4) in the next subsection.}

\rev{Lines 5 - 12, 2nd column, Page 12:  Just a general question spurred by the 
text here.  What criteria would be applied to determine which FPG is better 
than another?}

\rev{Line 57, 1st column, Page 13: change ``as a input'' to ``as an input''}

\rev{Now for the request of more significant importance.  This is coming from 
someone with significant experience in MDO and who uses a lot of the 
analysis tools you specified in your case study on a daily basis.  I really 
liked this article but felt it was lacking some punch in the results 
section.  This was namely in the form of quantitative results.  You vaguely 
give some in the form of the different resulting connectivity graphs and 
this was good but to really drive home the value of your research to a 
potential user I think you could do a bit better with very minimal work. 
The potential user of this methodology will be concerned with one of two 
things: 1) minimizing computational time or 2) increasing the accuracy of 
the results.  I think you've given examples of these two scenarios already 
in figures 15 and 16, so why not provide a table of the CPU times (1) and/or 
deviation from higher order analyses (2) for each of these graphs.  Doesn't 
have to necessarily be these two metrics but something to really tie the 
output of your methodology to something tangible for comparison purposes. 
You could even include the metrics for a randomly obtained FPG as a 
baseline.  Then the person reading the article can say, "Wow, given a set of 
analysis tools and a general problem, I can employ this method to obtain an 
FPG that significantly reduces my overall computational time." or "Wow, 
given a set of analysis tools and a general problem, I can employ this 
method to obtain an FPG that significantly improves my overall computational 
accuracy while keeping my computational efficiency as high as possible."}

### 2013-09-04 conclusions.tex
finished reviewer 1 comments

A collision corresponds to a situation in which two or more analysis tools compute separate instances of the same variable. 
To create a valid problem formulation, all collisions must be resolved either by removing redundant analysis tools or by introducing logic or solvers into the data flow to choose between or aggregate multiple instances during execution. 
The identification of collisions in an MCG is especially important for problem formulations involving multiples analysis tools of different fidelities; in this context, a collision represents an opportunity to formulate a multi-fidelity problem that can take best advantage of the efficiency/accuracy tradeoff between different analysis tools. 
Resolving collisions and filling holes in an MCG requires choices by the user to specify the overall MDAO problem that he/she desires to solve, i.e. to produce the required set of system-level outputs given a specified set of design variables. 
One benefit of a graph theoretic approach is the standard algorithms which can be used to inform the user's decisions, such as cycle detection, minimum spanning trees, and shortest path algorithms.

>>>>

A collision corresponds to a situation in which two or more analysis tools compute separate instances of the same variable. 
To create a valid problem formulation, all collisions must be resolved either by removing redundant analysis tools or by introducing logic or solvers into the data flow to choose between or aggregate multiple instances during execution. 
The identification of collisions in an MCG is especially important for problem formulations involving multiple analysis tools of different fidelities; in this context, a collision represents an opportunity to formulate a multi-fidelity problem that can take best advantage of the efficiency/accuracy tradeoff between different analysis tools. 
Resolving collisions and filling holes in an MCG requires choices by the user to specify the overall MDAO problem that he/she desires to solve, i.e. to produce the required set of system-level outputs given a specified set of design variables. 
One benefit of a graph theoretic approach is the standard algorithms which can be used to inform the user's decisions, such as cycle detection, minimum spanning trees, and shortest path algorithms.

### 2013-09-04 example_problem.tex
finished reviewer 1 comments

Step (B) is the production of the MCG.
	In this example, the MCG is formed by using a consistant variable naming convention and then connecting all variables with the same name with connection edges. 
	Table \ref{t:ins and outs} presents the full list of variables in the leftmost column and indicates whether the variable is an input or an output of each analysis tool. 
	Some variables, such as geometry and performance, represent groups of variables that are passed as arrays or other data structures due to their similarity. 
	This bundling of variables is not fundamental and does not limit the generality of this example; rather, it simplifies the presentation.
	\begin{table*}[htb!]
	  \centering
	  \caption{Analysis code input and output description}
		\begin{tabular}{ccccccccc}
		\toprule
		variable & \multicolumn{8}{c}{analysis code} \\
		\midrule
			  & VSP   & PDCYL & NPSS  & VORLAX & PMARC & WATE  & FLOPSa & FLOPSb \\
		geometry & in    & in    &       & in    & in    & in    & in    & in \\
		number of engines &       &       & in    &       &       &       & in    & in \\
		mission &       &       &       &       &       &       & in    & in \\
		fuselage weight &       & out   &       &       &       &       & in    & in \\
		wing weight &       & out   &       &       &       &       & in    & in \\
		engine weight &       &       &       &       &       & out   & out   & in \\
		wetted area & out   &       &       &       &       &       & in    & in \\
		inviscid drag &       &       &       & out   & out   &       &       & in \\
		drag  &       &       & in    &       &       &       & out   & out \\
		engine performance &       &       & out   &       &       & in    & out   & in \\
		performance &       &       &       &       &       &       & out   & out \\
		total weight &       & in    &       &       &       &       & out   & out \\
		\bottomrule
		\end{tabular}
	  \label{t:ins and outs}
	\end{table*}

@@

Step (D) now procededs by executing the FPG algorithm. 
An FPG cannot be obtained because the variable nodes representing the number of engines are holes for NPSS, FLOPSa, and FLOPSb, and the variable nodes representing the mission definition are holes for FLOPSa and FLOPSb. 
These holes are propagated upstream by the FPG algorithm and create holes in the objectives, thereby preventing a valid FPG.

@@

	Alternatively, it may be desirable to resolve collisions by preferentially choosing analysis tools with favorable properties, such as higher accuracy or minimal run time, whenever possible.
	One method to implement an FPG in this paradigm is to use a ranking system. Each analysis block is assigned a value by the designer describing how desireable it is for this block to be present in the FPG. The connection edges directed out of each analysis block are then assigned a weight equal to the value assigned to the analysis block. Finally, each collision is resolved by selecting the edges with the highest weights.
For the current example problem, consider the rankings for each analysis code given in Table \ref{t:rankings}.
	\begin{table}[htbp]
	  \centering
	  \caption{Example ranking of importance}
		\begin{tabular}{cc}
		\toprule
		analysis code & rank \\
		\midrule
		VSP   & 5 \\
		PDCYL & 5 \\
		NPSS  & 4 \\
		PMARC & 4 \\
		FLOPSb & 4 \\
		VORLAX & 3 \\
		WATE  & 3 \\
		FLOPSa & 2 \\
		\bottomrule
		\end{tabular}%
	  \label{t:rankings}%
	\end{table}%
	The resulting FPG has four cycles and is shown in Fig.~\ref{f:FPG highest rank}. 
	\begin{figure*}[htb!]
	  \begin{center}
		\includegraphics[width=6in]{images/FPG_edit_ranking}
	  \end{center}
	  \caption{FPG obtained using a ranking system.}
	\label{f:FPG highest rank}
	\end{figure*}

	Finally, consider a case in which 
the induced drag input into FLOPSb is a multi-fidelity input, meaning that 
 multiple analysis codes calculate the same variable. 
	This multi-fidelity formulation is implemented by setting the upper indegree limit for this node as two and then repreating the (automated) process. 
	In this case, both VORLAX and PMARC are retained, resulting in an FPG that omits only FLOPSa.

>>>>

Step (B) is the production of the MCG.
	In this example, the MCG is formed by using a consistent variable naming convention and then connecting all variables with the same name with connection edges. 
	Table \ref{t:ins and outs} presents the full list of variables in the leftmost column and indicates whether the variable is an input or an output of each analysis tool. 
	Some variables, such as geometry and performance, represent groups of variables that are passed as arrays or other data structures due to their similarity. 
	This bundling of variables is not fundamental and does not limit the generality of this example; rather, it simplifies the presentation.
	\begin{table*}[htb!]
	  \centering
	  \caption{Analysis code input and output description}
		\begin{tabular}{ccccccccc}
		\toprule
		variable & \multicolumn{8}{c}{analysis code} \\
		\midrule
			  & VSP   & PDCYL & NPSS  & VORLAX & PMARC & WATE  & FLOPSa & FLOPSb \\
		geometry & in    & in    &       & in    & in    & in    & in    & in \\
		number of engines &       &       & in    &       &       &       & in    & in \\
		mission &       &       &       &       &       &       & in    & in \\
		fuselage weight &       & out   &       &       &       &       & in    & in \\
		wing weight &       & out   &       &       &       &       & in    & in \\
		engine weight &       &       &       &       &       & out   & out   & in \\
		wetted area & out   &       &       &       &       &       & in    & in \\
		inviscid drag &       &       &       & out   & out   &       &       & in \\
		drag  &       &       & in    &       &       &       & out   & out \\
		engine performance &       &       & out   &       &       & in    & out   & in \\
		performance &       &       &       &       &       &       & out   & out \\
		total weight &       & in    &       &       &       &       & out   & out \\
		\bottomrule
		\end{tabular}
	  \label{t:ins and outs}
	\end{table*}

@@

Step (D) now proceeds by executing the FPG algorithm. 
An FPG cannot be obtained because the variable nodes representing the number of engines are holes for NPSS, FLOPSa, and FLOPSb, and the variable nodes representing the mission definition are holes for FLOPSa and FLOPSb. 
These holes are propagated upstream by the FPG algorithm and create holes in the objectives, thereby preventing a valid FPG.

@@

	Alternatively, it may be desirable to resolve collisions by preferentially choosing analysis tools with favorable properties, such as higher accuracy or minimal run time, whenever possible.
	One method to implement an FPG in this paradigm is to use a ranking system. Each analysis block is assigned a value by the designer describing how desirable it is for this block to be present in the FPG. The connection edges directed out of each analysis block are then assigned a weight equal to the value assigned to the analysis block. Finally, each collision is resolved by selecting the edges with the highest weights.
For the current example problem, consider the rankings for each analysis code given in Table \ref{t:rankings}.
	\begin{table}[htbp]
	  \centering
	  \caption{Example ranking of importance}
		\begin{tabular}{cc}
		\toprule
		analysis code & rank \\
		\midrule
		VSP   & 5 \\
		PDCYL & 5 \\
		NPSS  & 4 \\
		PMARC & 4 \\
		FLOPSb & 4 \\
		VORLAX & 3 \\
		WATE  & 3 \\
		FLOPSa & 2 \\
		\bottomrule
		\end{tabular}%
	  \label{t:rankings}%
	\end{table}%
	The resulting FPG has four cycles and is shown in Fig.~\ref{f:FPG highest rank}. 
	\begin{figure*}[htb!]
	  \begin{center}
		\includegraphics[width=6in]{images/FPG_edit_ranking}
	  \end{center}
	  \caption{FPG obtained using a ranking system.}
	\label{f:FPG highest rank}
	\end{figure*}

	Finally, consider a case in which 
the induced drag input into FLOPSb is a multi-fidelity input, meaning that 
 multiple analysis codes calculate the same variable. 
	This multi-fidelity formulation is implemented by setting the upper indegree limit for this node as two and then repeating the (automated) process. 
	In this case, both VORLAX and PMARC are retained, resulting in an FPG that omits only FLOPSa.

### 2013-09-04 intro.tex
finished reviewer 1 comments

    The time and expense for setting up 
    analysis models is growing in conjunction with the increasing scope of MDO problems. 
    Multidisciplinary Design Analysis and Optimization (MDAO)
    frameworks such as OpenMDAO \cite{Gray2012} and {ModelCenter}\textregistered have enabled a new level of analysis tool integration 
    and have paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. This 
    new capability has created a new challenge. Even models with tens of analyses could include hundreds or thousands
    of variables that are interdependent and must be linked in the framework. 
    A common situation is that different analyses provide differing values for the 
    same physical quantity, often in different data formats, and these conflicts need to be resolved. 
    These occurrences are particularly acute in situations in which analysis tools have differing fidelities. For
    example, an abstracted aerodynamic analysis such as an empirical drag buildup model may return only
    integrated drag, whereas a CFD tool may return pressure and shear stress distributions across the entire
    surface grid. If an analysis downstream of the aerodynamics tool needs only integrated drag as an input,
    then the designer has a free choice of which of the two possible aerodynamics analysis tools to select to
    provide the drag estimate (presuming that drag can be computed from surface distributions by a simple
    integration algorithm). On the other hand, if a downstream analysis needs a pressure distribution in
    order to compute pitching moment, for example, then any feasible MDO problem formulation must
    include CFD or similar analysis in the data flow, regardless of whether the empirical drag buildup model
    is also included.

@@

>>>>

    The time and expense for setting up 
    analysis models is growing in conjunction with the increasing scope of MDO problems. 
    Multidisciplinary Design Analysis and Optimization (MDAO)
    frameworks such as {ModelCenter}\textregistered, iSigt\textregistered, ModeFrontier\textregistered, and OpenMDAO \cite{Gray2012}and have enabled a new level of analysis tool integration 
    and have paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. This 
    new capability has created a new challenge. Even models with tens of analyses could include hundreds or thousands
    of variables that are interdependent and must be linked in the framework. 
    A common situation is that different analyses provide differing values for the 
    same physical quantity, often in different data formats, and these conflicts need to be resolved. 
    These occurrences are particularly acute in situations in which analysis tools have differing fidelities. For
    example, an abstracted aerodynamic analysis such as an empirical drag buildup model may return only
    integrated drag, whereas a CFD tool may return pressure and shear stress distributions across the entire
    surface grid. If an analysis downstream of the aerodynamics tool needs only integrated drag as an input,
    then the designer has a free choice of which of the two possible aerodynamics analysis tools to select to
    provide the drag estimate (presuming that drag can be computed from surface distributions by a simple
    integration algorithm). On the other hand, if a downstream analysis needs a pressure distribution in
    order to compute pitching moment, for example, then any feasible MDO problem formulation must
    include CFD or similar analysis in the data flow, regardless of whether the empirical drag buildup model
    is also included.

@@

\subsection{Potential Applications}

The syntax proposed here for defining graphs is partially motivated by the unique needs
of working with very large problems that have thousands of design variables. Such problems are 
large enough that many traditional design methods stumble simply because the problems 
are so large that one person can not maintain a complete view of the problem. Hence as problems 
grow to that size, software tools become needed to manage the complexity of the problem itself.

Even for small problems, algorithmic exploration of problem structure can yield deeper understanding 
of problems and more effective solution methods, as demonstrated by the work with the FDT. The syntax 
described here along with the rules for applying that syntax to the three phases of the design 
process provides a strong foundation to begin developing new algorithms for analyzing and modifying 
problem structure. This foundation makes it much easier to incorporate a standard view of 
problem definition into MDAO frameworks and by extension to implement any new algorithms inside 
those frameworks.

### 2013-09-04 mcg_fpg.tex
finished reviewer 1 comments

\subsection{Problem Solution Graph}
Once a suitable FPG has been obtained the issue of how to structure a solution strategy 
can be addressed. Similar to the way an MCG can be manipulated into a FPG, an FPG can be
further manipulated into a PSG. To get from an MCG to an FPG you eliminated analysis blocks and
connection edges to reduce the problem to something solvable, but to get from an FPG to a PSG you 
add in driver blocks, driven edges to describe the structure of the solution strategy being applied 
to the problem. For example, Figure \ref{f:selllar_psg} shows the PSG that solves the Sellar problem 
with an MDF solution strategy. There are two driver blocks in this PSG, one for the solver that 
ensures compatibility and another for the optimizer that picks the design variables.

\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=3.0in]{images/sellar_psg}
    \end{center}
    \vspace{-10pt}
    \caption{Example PSG for solving the Sellar problem with MDF strategy}
    \label{f:sellar_psg}
\end{figure}

The PSG in figure \ref{f:selllar_psg} can be compared to the XDSM for the same situation 
in Fig. \ref{fig:XDSM_simple}. While the two diagrams contain much of the same information 
there is a significant difference. The PSG describes the data passing that happens between 
the drivers and the rest of the problem, but it does not describe when that data passing 
will occur (the execution order of the blocks). In the proposed graph syntax edges describe 
the passing of data between one node an another, but they do not define when that 
communication happens.

>>>>

\subsection{Problem Solution Graph}
Once a suitable FPG has been obtained the issue of how to structure a solution strategy 
can be addressed. Similar to the way an MCG can be manipulated into a FPG, an FPG can be
further manipulated into a PSG. To get from an MCG to an FPG you eliminated analysis blocks and
connection edges to reduce the problem to something solvable, but to get from an FPG to a PSG you 
add in driver blocks, driven edges to describe the structure of the solution strategy being applied 
to the problem. For example, Figure \ref{f:sellar_psg} shows the PSG that solves the Sellar problem 
with an MDF solution strategy. There are two driver blocks in this PSG, one for the solver that 
ensures compatibility and another for the optimizer that picks the design variables.

\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=3.0in]{images/sellar_mdf_psg}
    \end{center}
    \vspace{-10pt}
    \caption{Example PSG for solving the Sellar problem with MDF strategy}
    \label{f:sellar_psg}
\end{figure}

The PSG in figure \ref{f:sellar_psg} can be compared to the XDSM for the same situation 
in Fig. \ref{fig:XDSM_simple}. While the two diagrams contain much of the same information 
there is a significant difference. The PSG describes the data passing that happens between 
the drivers and the rest of the problem, but it does not describe when that data passing 
will occur (the execution order of the blocks). In the proposed graph syntax edges describe 
the passing of data between one node an another, but they do not define when that 
communication happens.
